% Type your Acknowledgements below. Delete all the text after the double-spacing command.
\chapter{Acknowledgements}
\doublespacing

I'd like to extend my gratitude to my technical advisor, Dr. Chris Hain, for his guidance and dedication toward me throughout the process of developing this work. This has been a long process with plenty of setbacks and diversions, and I am thankful for the patience, gentle correction, and freedom he and the rest of my committee has afforded me. My appreciation also goes to Dr. Sundar Christopher for his kindness in providing both scientific and administrative direction, and to Dr. Sean Freeman for engaging with interest and providing insight on the technical details of my project. Thanks also to Ryan Wade for being a protean beacon of knowledge and experience, and for offering me a wealth of advice and many opportunities, and to Paul Meyer for his close mentorship and many hours of conversation during my undergrad years.

Naturally, this wouldn't be possible without the help of a community of people around me. It has been a pleasure to forge strong bonds with my fellow graduate students as we overcome numerous challenges together. I also wouldn't have had the fortitude to finish this effort if it weren't for weekend escapes into caves and over mountains with my dear friends and my partner Shae, and the mental respite they provide. The entire institution of the National Space Science and Technology Center -- the SPoRT team in particular -- also has my appreciation and esteem; it is an honor and a source of constant motivation to work alongside such a diverse and passionate group of experts.

Finally, my greatest thanks goes to my family, who have continued to support and believe in me, even when they aren't sure where all my time is going. Without their confidence, love, and prayers, this process could not have even begun.
